On the Functionality of the Multilayer Perceptron

### 2020-12-23 Thesis/Drafts/Buell-PHYS799-ThesisDraft4.tex
Organized Final Draft

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being in many cases, but is far more difficult for a computer to do. In order to automate audio source identification, we have designed and constructed a multimodal neural network that can classify audio files by what musical instrument the sound within it most resembles. The network uses two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate and transform the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become a topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place information into distinct categories called \textit{classes} based on inherent properties of the data. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do without the aide of some form of automation.

@@

\paragraph*{}The design of our neural network function combines two distinct architecture types by using two distinct entry arrays which are processed, and a single output is formed. This type of neural network is referred to as a \textit{multimodal} neural network \cite{Li,Ngiam}. Our model contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computational graph that is designed to process it's respective inputs types. The activations at the end of each layer are concatenated into a new layer, which is further processed and then used to generate a single prediction. This output that is aggregated from two inputs shows on average, superior classification performance when compared to either of it's constituent unimodal architectures alone.

@@

\paragraph*{}To ensure that these data sets are formatted consistently, we have designed a MATLAB program to read each audio file from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}). After the samples are formatted, we have produced a Python script which iterates through each new file and assigns a class label based on the name of the audio file. This has been design to account for the specific naming cases that may arise. We assemble a dictionary-like structure of all training files with organizes the local file location of each sample and provides an integer class label and string class label for each sample.

@@

\paragraph*{}Below we outline a standard "forward pass" algorithm for a general feed-forward deep neural network. We presume each layer to be a node in a computational graph that contains a method "\textit{call}" which executes the layer's main function as in Eq. (\ref{eqn-altLayerFunction}). Additionally, each node contains a pointer "\textit{next}" which gives the next layer in the network chain. Each layer should also store it's linear activations, $z^{(l)}$ and it's output activations $x^{(l)}$ for training purposes. This algorithm is adapted from Goodfellow, \cite{Goodfellow}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used for image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. We call the shape the \textit{convolution kernel size}, and often choose $m = n$ \cite{Loy, Goodfellow}. The step size in each of the two dimensions is known as the \textit{stride size}, which we choose to be $1$ in both dimensions.

@@

\paragraph*{}For 2D convolution layer $(l)$, given input $x^{(l-1)}$, we compute the activations of the $k$-th filter $x^{(l)}_{k}$ using feature map $W^{(l)}_k$ and bias neuron $b^{(l)}_k$ as \cite{Goodfellow}:
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ b^{(l)}_k +  \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}_k(i,j) x^{(l-1)}_k(i - u,j - v) \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

@@

\paragraph*{}A neural network's purpose is to produce a function $F^*$ that approximates an unknown function $F$, using a set of parameters, $\Theta$. The model must have a procedure for updating the parameters in $\Theta$ to allow for a reasonably close approximation of $F$ \cite{Goodfellow}. To better understand this, we turn to Tom Mitchell's explanation of a learning algorithm \cite{Mitchell}:
\begin{quote}
A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if it's performance at tasks in $T$, as measured by $P$, improves with experience $E$.
\end{quote}
Without any direct human intervention, a model must update itself to improve it's performance at a give task as new information is presented to it. To do this, the model must be constructed with a training procedure in mind.

@@

\paragraph*{}Suppose we pass a training sample into the neural network. This sample is represented by the feature-vector $x^{(0)}$, with an expected outcome given by the one-hot-encoded vector $y$.  After the network finishes processing, it's output activations, given by the vector $x^{(L-1)}$, also noted as $y^*$, represents the prediction for the class label. For a reasonably trained model, we expect the vector $y^*$ to be "similar" to $y$, indicating that model has made a a reasonable prediction with the given input values. Alternatively, for an untrained network, $y^*$ is not likely to be "similar" to $y$ at all, indicating that the model has made a poor prediction with the given input.

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to perform the optimization task to reduce the cost value after reciveing each new training sample. Regression models may often used a \textit{mean-squared error} cost function which can often be solved analytically, while high dimensional neural networks, such as the one used in this project, outlined in section (\ref{subsec-Architecture}), can only up updated iteratively \cite{Goodfellow,James,Loy}. We use a variation of gradient descent much like the one in Eq. (\ref{eqn-GradientLearning}). A standard stochastic gradient-based or batch gradient descent learning algorithm can be fast, but are often prone errors such as vanishing or exploding gradients \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an optimizer that is more stable, and robust in it's ability to drive the cost function to a successively lower value.

@@

\paragraph*{}In addition to choosing a strong set of features, and an appropriate optimizer for a classification task, it is also important to combine the predictors with a complementary network architecture. Since a neural network is a computational graph with each layer represented by a node, it can take on a nearly infinite number of different shapes and structures \cite{Goodfellow,Virtanen}. The structure of the network defines the composition of it's function, $F^*$ as in Eq.(\ref{eqn-FunctionGraph}), and thus affects its performance in completing any given task \cite{Geron}. The layer structure of the neural network is referred to as its \textit{architecture}, and define the networks \textit{hypothesis - space} \cite{Goodfellow}.

@@

\paragraph*{}As discussed in section (\ref{sec-TheNeuralNetwork}), performance quality of a neural network is greatly dependent on the properties of the chosen features \cite{Virtanen,Liu}. To ensure that the classifier model performs adequately and consistently, we must choose these features to have high variance between classes, and low variance within each class. This enables a neural network to development clear decision boundaries between each unique class \cite{James,Serizel}. To develop this set of features, we explore the physical and mechanical properties of musical instruments and the sounds that they create.

\paragraph*{}Since computer algorithms such are neural networks are built to handle exclusively quantitative information, we must ensure that each property can be represented by a numerical quantity \cite{Geron,Levine}. Where a human would use qualitative adjectives such as \textit{bright} or \textit{harsh} or \textit{percussive} to describe sound, computers require a numerical descriptor instead. The goal of this section is to develop the attributes of sound that enable us to describe a sound wave in a quantitative and compact way. Thus we use values that can be expressed as numbers such as \textit{amplitude envelopes}, \textit{frequency spectra}, or \textit{formant structure}. For this section, we represent waveforms as \textit{spectrograms} to capture the time-space and frequency-space development of energy simultaneously. We motivate and describe the production of a spectrogram in section (\ref{subsec-Spectrogram}).

@@

\paragraph*{}Idiophones in particular can be characterized by their \textit{transient response} \cite{White}. The transient response is the time-evolution response of the waveform in time, which we normally break into \textit{attack}, \textit{decay}, \textit{sustain}, and \textit{release}. For most idiophones, the system is struck once and left to a free or dampened vibration; which we can mathematically model as a rigid body reacting to an impulse \cite{Hunter,Olson}.

@@

\paragraph*{}The wave equation is a \textit{linear} partial differential equation which allows for a linear combination of solutions to form a sound wave. In the case of the taught finite string that is fixed at both ends, such a guitar, viola or cello, the a solution tends to take the form of a sine or cosine function \cite{Hunter, Taylor, White}. Notice the terms arise by the many a parallel yellow and orange lines in each spectrogram in Fig. (\ref{fig-PropertiesChordophones}). The verical position of any line is given by the linear frequency of the sinusoid, and the brightness is proportional to the coefficient of the combination. The boundary conditions, and most forms of standard initial conditions allow for the each sinusoidal to fit exactly an integer or half-integer number wavelengths into the region $0 \leq x \leq L$ \cite{Haberman,Hunter}. Players of chordophones change pitch by shortening the length of the vibrating string or applying additional tension to the string. This changes the length of portion of the string that vibrates, and causes the production of a new pitch.

\paragraph*{}Each chordophone instrument does produce a slightly different frequency spectrum, hence each provides a slightly different timbre to the listener, so differentiating between them is possible. Because of this, we find that the spectrogram representation of the waveform is so critical to classification success because the convolution layers allow for the detection of small variations in the frequency-time domain where humans would otherwise be unable to \cite{Loy,Virtanen}. However, the size of the instrument limits the length of the string that vibrates which leads to limitations in the range of the overtones the instruments will produce. For example, a cello playing a low $C1$ will have overtones extending up into the $12$ kHz range, but the energy will be very low there, thus have a very low frequency center of mass. Compare this with a violin playing a $C6$, will have overtones into the $20$ kHz range, and will produce a very high frequency center of mass. While the ranges of the musical instruments overlap a great deal, the energy distribution in frequency space can be treated as a 1D mass distribution, and the frequency center of masses of instruments with similar timbres can still differ greatly \cite{Serizel}.

@@

\paragraph*{}An aerophone is a musical instrument that produces sounds through the vibration of an air molecules in a column-like structure. This is the largest category of instruments that we have, and is commonly broken down into smaller subcategories such as \textit{woodwind} and \textit{brass} instruments. From our data set - alto flutes, alto saxophones, bass clarinets, bass flutes, bassoons, $B\flat$ clarinets, contrabassoons, $E\flat$ clarinets, English horns, flutes, oboes, tenor saxophones, and soprano saxophones are all examples of woodwind aerophones. Additionally, bass trombones, french horns, tenor trombones, trombones, $B\flat$ trumpets, and tubas are all examples of brass aerophones \cite{Hornbostel,Olson,White}.

@@

\paragraph*{}We group sounds produced from non-physical musical instruments in this final category because they are not considered in the original Hornbostel-Sachs system. From our data set, this category includes the four simplest waveforms begin, sine waves, sawtooth waves, square waves, and triangle waves, as well as white colored noise \cite{White}. These sounds have been produced synthetically through a MATLAB program specifically for this project. Because of their synthetic nature, they do not display many of the physical or mechanical properties of the previous sections.

\paragraph*{}Each of the five synthetically generated waveform types were produce to have a constant envelope with all time. This means that each waveform has no characteristic amplitude shape and the spectrogram shows no time-evolution of energy as expected. Each of the the signals can be composed of a linear combinations of oscillating sinusoidal at various different frequencies and energies. For example, a sine wave consists only of a single oscillating term, while a square wave consists of repeated odd harmonics. White noise consists of a roughly uniform distribution of energy, thus shows no energy evolution in time or frequency space. These behaviors can be observed in Fig. (\ref{fig-PropertiesSynths}).

@@

\paragraph*{}Classification tasks require a set of inputs called \textit{predictors} or \textit{features} \cite{James,Loy,Serizel}. A feature is a quantitative, low-dimensional representation of a sample that conveys its important characteristics. For example, in classifying types of animals, a feature could be the mass, volume or number of teeth on the animal. Typically, we produce a set of $p$ features for each sample and assemble them into a $p$-dimensional vector called a \textit{feature vector} which acts as a list of descriptive qualities of the sample \cite{Geron,James}. This list is presented to the neural network and used to 
Audio machine learning researchers M. Kashif Saeed Khan and Wasfi G. Al-Khatib describe the vitality of feature selection \cite{Khan}:
\begin{quote}
The data reduction stage which is also called feature extraction, consists of discovering a few important facts about each class. The choice of features is critical as it greatly affects the accuracy of audio classification. 
\end{quote}
An adequate selection of features is vital to the process of training and evaluation of the classifier model \cite{Mierswa,Serizel,Liu}. Consider the task of identifying fish from sharks using only the presence of gills. Since both sharks and fish have gills, the features are correct representations, but not useful for discerning them. A more appropriate choice would be compare the number of gills that fish or sharks have, which would provide a correct representation and allows for differentiation.

\paragraph*{}In the biological process of categorizing sound sources, the time-domain waveform is usually enough to match the sound to a source \cite{Olson}. However, a computer representation of a waveform is an array-like structure of values sampled discretely in time and volume \cite{Virtanen,Liu}. Presenting the raw time-series waveform to a model directly has experimentally shown to be unstable in for classification tasks so we must develop a set of features that can describe important properties of the waveform in a far more efficient manner \cite{Goodfellow,James,Serizel}. To ensure the construction of a suitable model, we derive features based from three representations of the audio file: (i) a spectrogram matrix, (ii) the time-space, and (iii) the frequency-space. It is important to note that although this algorithm will map audio files to source instruments, it will never actually be presented with a waveform directly, and instead will rely solely on the features to make predictions.

@@

\paragraph*{}From each audio file sample, we calculate the values for the same predictors. The predictors are arranged into two separated structures, each called a \textit{design matrix}, described in section (\ref{subsubsec-DesignMatrix}) \cite{Goodfellow,Geron}. For a model with $p$ predictors, feature space is a $p$-dimensional vector space where each basis vector represents a predictor \cite{Goodfellow,James}. Each sample can be described by a $p$-dimensional vector, given by each row of the design matrix $X_2$. For samples $x^{(a)}$ and $x^{(b)}$ that belong to the same class, we expect the vector to be similar, $x^{(a)} \approx x^{(b)}$ in some way. This indicates that the two samples lie "close together" in space. Similarly, for samples $x^{(c)}$ and $x^{(d)}$ that belong to different classes, we expect the vectors to be very different, indicating that the two sample lie "far apart" in space. Note that this is an idealization, and not always the defining case for the decision function behavior.

\paragraph*{}We can expand this idea by considering by how features should behave within each class, and between each class. Tuomas Virtanen, machine learning and audio engineer writes in his book, "Computational Analysis of Sound Scene and Events" about specifically how features should be selected \cite{Virtanen}:
\begin{quote}
For recognition algorithms, the necessary property of the acoustic features is low variability among features extracted from examples assigned to the same class, and at the same time high variability allowing distinction between features extracted from examples assigned to different classes.
\end{quote}
This is to say, features should \textit{ideally} form non-overlapping groups or bubbles that allows classifiers to effectively differentiate categories. Given the complexity of real-world problems, this is an often unobtainable \cite{Goodfellow,James}. Under the right circumstances, classifiers can still produce parameters that allow for reasonable performance \cite{Geron,Loy}. In practice, a neural network does not rely on comparing the norms of different samples in feature space, but rather uses something called a \textit{decision boundary}.

@@

\paragraph*{}A trained classifier makes a prediction in the form of a statistical likelihood of a sample belonging to a particular class \cite{Goodfellow,James}. A $k$- categories classifier contains $k$ output neurons, each one representing a different class. When a set of features is presented to the classifier, the value contained within each neuron is the "confidence" on the interval $[0,1]$ that a sample belongs to that class. For example, if neuron $q$ has an activation value of $0.25$, we say that there is a $25\%$ chance that the sample belongs to class $q$. Most commonly, we take the neuron with the highest activation value to be the chosen class, even if the next highest activation is a close second \cite{Loy}. Thus the formal prediction function is the index of the maximum value in the output vector.

\paragraph*{}This concept stems from the idea that when a trained neural network classifier makes a prediction, it is basing class probabilities on a set of generated \textit{decision boundaries} \cite{Geron,James}. To explore decision boundaries, consider a simplified toy data set with a 2D feature space for a classifier that must only discern between \textit{Class A} and \textit{Class B}. We can visualize each sample in this 2D space, and color-code each sample according according to it's class label as in Fig. (\ref{fig-LinSep1}).

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.35]{../MakeFigures/LinearNonSeparable}
\end{center}
\caption{Feature Space with near linear non-separability between two classes}
\label{fig-LinNonSep}
\end{figure}

@@

>>>>

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being in many cases, but is far more difficult for a computer to do. In order to automate audio source identification, we have designed and constructed a multimodal neural network that can classify audio files by what musical instrument the sound within it most resembles. The network uses two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate and transform the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of its constituent unimodal architectures. Training data for this experiment is from studio recordings of the London Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become a topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals exist all over the modern world \cite{Virtanen}. In each context, the general idea is to place information into distinct categories called \textit{classes} based on inherent properties of the data. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do without the aide of some form of automation.

@@

\paragraph*{}The design of our neural network function combines two distinct architecture types by using two distinct entry arrays which are processed, and a single output is formed. This type of neural network is referred to as a \textit{multimodal} neural network \cite{Li,Ngiam}. Our model contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is its own computational graph that is designed to process its respective inputs types. The activations at the end of each layer are concatenated into a new layer, which is further processed and then used to generate a single prediction. This output that is aggregated from two inputs shows on average, superior classification performance when compared to either of its constituent unimodal architectures alone.

@@

\paragraph*{}To ensure that these data sets are formatted consistently, we have designed a MATLAB program to read each audio file from its original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}). After the samples are formatted, we have produced a Python script which iterates through each new file and assigns a class label based on the name of the audio file. This has been design to account for the specific naming cases that may arise. We assemble a dictionary-like structure of all training files with organizes the local file location of each sample and provides an integer class label and string class label for each sample.

@@

\paragraph*{}Below we outline a standard "forward pass" algorithm for a general feed-forward deep neural network. We presume each layer to be a node in a computational graph that contains a method "\textit{call}" which executes the layer's main function as in Eq. (\ref{eqn-altLayerFunction}). Additionally, each node contains a pointer "\textit{next}" which gives the next layer in the network chain. Each layer should also store its linear activations, $z^{(l)}$ and its output activations $x^{(l)}$ for training purposes. This algorithm is adapted from Goodfellow, \cite{Goodfellow}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used for image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's its name from the implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. We call the shape the \textit{convolution kernel size}, and often choose $m = n$ \cite{Loy, Goodfellow}. The step size in each of the two dimensions is known as the \textit{stride size}, which we choose to be $1$ in both dimensions.

@@

\paragraph*{}For 2D convolution layer $(l)$, given input $x^{(l-1)}$, we compute the activations of the $k$-th filter $x^{(l)}_{k}$ using feature map $W^{(l)}_k$ and bias neuron $b^{(l)}_k$ as \cite{Goodfellow}:
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ b^{(l)}_k +  \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}_k(i,j) x^{(l-1)}_k(i - u,j - v) \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has its own $n \times m$ weighting matrix and appropriately shaped bias .

@@

\paragraph*{}A neural network's purpose is to produce a function $F^*$ that approximates an unknown function $F$, using a set of parameters, $\Theta$. The model must have a procedure for updating the parameters in $\Theta$ to allow for a reasonably close approximation of $F$ \cite{Goodfellow}. To better understand this, we turn to Tom Mitchell's explanation of a learning algorithm \cite{Mitchell}:
\begin{quote}
A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if its performance at tasks in $T$, as measured by $P$, improves with experience $E$.
\end{quote}
Without any direct human intervention, a model must update itself to improve its performance at a give task as new information is presented to it. To do this, the model must be constructed with a training procedure in mind.

@@

\paragraph*{}Suppose we pass a training sample into the neural network. This sample is represented by the feature-vector $x^{(0)}$, with an expected outcome given by the one-hot-encoded vector $y$.  After the network finishes processing, its output activations, given by the vector $x^{(L-1)}$, also noted as $y^*$, represents the prediction for the class label. For a reasonably trained model, we expect the vector $y^*$ to be "similar" to $y$, indicating that model has made a a reasonable prediction with the given input values. Alternatively, for an untrained network, $y^*$ is not likely to be "similar" to $y$ at all, indicating that the model has made a poor prediction with the given input.

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to perform the optimization task to reduce the cost value after reciveing each new training sample. Regression models may often used a \textit{mean-squared error} cost function which can often be solved analytically, while high dimensional neural networks, such as the one used in this project, outlined in section (\ref{subsec-Architecture}), can only up updated iteratively \cite{Goodfellow,James,Loy}. We use a variation of gradient descent much like the one in Eq. (\ref{eqn-GradientLearning}). A standard stochastic gradient-based or batch gradient descent learning algorithm can be fast, but are often prone errors such as vanishing or exploding gradients \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an optimizer that is more stable, and robust in its ability to drive the cost function to a successively lower value.

@@

\paragraph*{}In addition to choosing a strong set of features, and an appropriate optimizer for a classification task, it is also important to combine the predictors with a complementary network architecture. Since a neural network is a computational graph with each layer represented by a node, it can take on a nearly infinite number of different shapes and structures \cite{Goodfellow,Virtanen}. The structure of the network defines the composition of its function, $F^*$ as in Eq.(\ref{eqn-FunctionGraph}), and thus affects its performance in completing any given task \cite{Geron}. The layer structure of the neural network is referred to as its \textit{architecture}, and define the networks \textit{hypothesis - space} \cite{Goodfellow}.

@@

\paragraph*{}As discussed in section (\ref{sec-TheNeuralNetwork}), performance quality of a neural network is greatly dependent on the properties of the chosen features \cite{Virtanen,Liu}. To ensure that the classifier model performs adequately and consistently, we must choose these features to have high variance between classes, and low variance within each class. This enables a neural network to develop clear decision boundaries between each unique class \cite{James,Serizel}. To develop this set of features, we explore the physical and mechanical properties of musical instruments and the sounds that they create.

\paragraph*{}Since computer algorithms such as neural networks are built to handle mostly quantitative information, we must ensure that each property can be represented by a numerical quantity \cite{Geron,Levine}. Where a human would use qualitative adjectives such as \textit{bright} or \textit{harsh} or \textit{percussive} to describe sound, computers require a numerical descriptor instead. The goal of this section is to develop the attributes of sound that enable us to describe a sound wave in a quantitative and compact way. Thus we use values that can be expressed as numbers such as \textit{amplitude envelopes}, \textit{frequency spectra}, or \textit{formant structure}. For this section, we represent waveforms as \textit{spectrograms} to capture the time-space and frequency-space development of energy simultaneously. We motivate and describe the production of a spectrogram in section (\ref{subsec-Spectrogram}).

@@

\paragraph*{}Idiophones in particular can be characterized by their \textit{transient response} \cite{White}. The transient response is the time-evolution response of the waveform, which we normally break into \textit{attack}, \textit{decay}, \textit{sustain}, and \textit{release}. For most idiophones, the system is struck once and left to a free or dampened vibration; which we can mathematically model as a rigid body reacting to an impulse \cite{Hunter,Olson}.

@@

\paragraph*{}The wave equation is a \textit{linear} partial differential equation which allows for a linear combination of solutions to form a sound wave. In the case of the taught finite string that is fixed at both ends, such a guitar, viola or cello, the a solution tends to take the form of a linear combination of sine or cosine function \cite{Hunter, Taylor, White}. Notice the terms arise by the many a parallel yellow and orange lines in each spectrogram in Fig. (\ref{fig-PropertiesChordophones}). The vertical position of any line is given by the linear frequency of the sinusoid, and the brightness is proportional to the coefficient of the combination. The boundary conditions, and most forms of standard initial conditions allow for the each sinusoidal to fit exactly an integer number wavelengths into the region $0 \leq x \leq L$ \cite{Haberman,Hunter}. Players of chordophones change pitch by shortening the length of the vibrating string or applying additional tension to the string. This changes the length of portion of the string that vibrates, and causes the production of a new pitch.

\paragraph*{}Each chordophone instrument does produces a slightly different frequency spectrum, hence each provides a slightly different timbre to the listener, so differentiating between them is possible but can be challenging. Because of this, we find that the spectrogram representation of the waveform is so critical to classification success because the convolution layers allow for the detection of small variations in the frequency-time domain where humans would otherwise be unable to \cite{Loy,Virtanen}. However, the size of the instrument limits the length of the string that vibrates which leads to limitations in the range of the overtones the instruments will produce. For example, a cello playing a low $C1$ will have overtones extending up into the $12$ kHz range, but the energy will be very low there, thus have a very low frequency center of mass. Compare this with a violin playing a $C6$, will have overtones into the $20$ kHz range, and will produce a very high frequency center of mass. While the ranges of the musical instruments overlap a great deal, the energy distribution in frequency space can be treated as a 1D mass distribution, and the frequency center of masses of instruments with similar timbres can still differ greatly \cite{Serizel}.

@@

\paragraph*{}An aerophone is a musical instrument that produces sounds through the vibration of air molecules in a column-like structure. This is the largest category of instruments that we have, and is commonly broken down into smaller subcategories such as \textit{woodwind} and \textit{brass} instruments. From our data set - alto flutes, alto saxophones, bass clarinets, bass flutes, bassoons, $B\flat$ clarinets, contrabassoons, $E\flat$ clarinets, English horns, flutes, oboes, tenor saxophones, and soprano saxophones are all examples of woodwind aerophones. Additionally, bass trombones, french horns, tenor trombones, trombones, $B\flat$ trumpets, and tubas are all examples of brass aerophones \cite{Hornbostel,Olson,White}.

@@

\paragraph*{}We group sounds produced from non-physical musical instruments in this final category because they are not considered in the original Hornbostel-Sachs system. From our data set, this category includes the four simplest waveforms being, sine waves, sawtooth waves, square waves, and triangle waves, as well as white colored noise \cite{White}. These sounds have been produced synthetically through a MATLAB program specifically for this project. Because of their synthetic nature, they do not display many of the physical or mechanical properties of the previous sections.

\paragraph*{}Each of the five synthetically generated waveform types were produced to have a constant amplitude envelope with all time. This means that each waveform has no characteristic amplitude shape and the spectrogram shows no time-evolution of energy as expected. Each of the the signals can be composed of a linear combinations of oscillating sinusoidal at various different frequencies and energies. For example, a sine wave consists only of a single oscillating term, while a square wave consists of repeated odd harmonics. White noise consists of a roughly uniform distribution of energy, thus shows no energy evolution in time or frequency space. These behaviors can be observed in Fig. (\ref{fig-PropertiesSynths}).

@@

\paragraph*{}Classification tasks require a set of inputs called \textit{predictors} or \textit{features}. A feature is a quantitative, low-dimensional representation of a sample that conveys its important characteristics \cite{James,Loy,Virtanen}. For example, in classifying types of animals, a feature could be the mass, volume or number of teeth on the animal. Typically, we produce a set of $p$ features for each sample and assemble them into a $p$-dimensional vector called a \textit{feature vector} which acts as a list of descriptive qualities of the sample \cite{Geron,James}. This list is presented to the neural network and is transformed through the layers to produce an output. Audio machine learning researchers M. Kashif Saeed Khan and Wasfi G. Al-Khatib describe the vitality of feature selection \cite{Khan}:
\begin{quote}
The data reduction stage which is also called feature extraction, consists of discovering a few important facts about each class. The choice of features is critical as it greatly affects the accuracy of audio classification. 
\end{quote}
An adequate selection of features is vital to the process of training and evaluation of the classifier model \cite{Mierswa,Serizel,Liu}. Consider the task of identifying fish from sharks using only the presence of gills. Since both sharks and fish have gills, the features are correct representations, but not useful for discerning them. A more appropriate choice would be compare the number of gills that fish or sharks have, which would provide a correct representation and allows for differentiation.

\paragraph*{}In the biological process of categorizing sound sources, the time-domain waveform is usually enough to match the sound to a source \cite{Olson}. However, a computer representation of a waveform is an array-like structure of values sampled discretely in time and amplitude \cite{Virtanen,Liu}. Presenting the raw time-series waveform to a model directly has experimentally shown to be unstable for classification tasks so we must develop a set of features that can describe important properties of the waveform more efficiently \cite{Goodfellow,James,Serizel}. To ensure the construction of a suitable model, we derive features from three representations of the audio file: (i) a spectrogram matrix, (ii) time-space, (iii) the frequency-space. It is important to note that although this algorithm will map audio files to source instruments, it will never actually be presented with a waveform directly, and instead will rely solely on the features to make predictions.

@@

\paragraph*{}From each audio file sample, we calculate the values for the same predictors. The predictors are arranged into two separated structures, each called a \textit{design matrix}, described in section (\ref{subsubsec-DesignMatrix}) \cite{Goodfellow,Geron}. For a model with $p$ predictors, feature space is a $p$-dimensional vector space where each basis vector represents a predictor \cite{Goodfellow,James}. Each sample can be described by a $p$-dimensional vector, given by each row of the design matrix $X_2$. For samples $x^{(a)}$ and $x^{(b)}$ that belong to the same class, we expect the vector to be similar, $x^{(a)} \approx x^{(b)}$ in some way. This indicates that the two samples have similar properties in this space. Alternatively, for samples $x^{(c)}$ and $x^{(d)}$ that belong to different classes, we expect the vectors to be very different, indicating that the two samples have very different properties in this space. Note that this is an idealization, and not always the defining case for the decision function behavior.

\paragraph*{}We can expand this idea by considering by how features should behave within each class, and between each class. Tuomas Virtanen, machine learning and audio engineer writes in his book, "Computational Analysis of Sound Scene and Events" about specifically how features should be selected \cite{Virtanen}:
\begin{quote}
For recognition algorithms, the necessary property of the acoustic features is low variability among features extracted from examples assigned to the same class, and at the same time high variability allowing distinction between features extracted from examples assigned to different classes.
\end{quote}
This is to say, features should \textit{ideally} form non-overlapping groups or bubbles that allows the classifier model to effectively differentiate categories. Given the complexity of real-world problems, this is an often unobtainable \cite{Goodfellow,James}. Under the right circumstances, classifiers can still produce parameters that allow for reasonable performance \cite{Geron,Loy}. In practice, a neural network does not rely on comparing the vectors directly for classification, but rather uses something called a \textit{decision boundary}.

% ===============================

@@

\paragraph*{}A trained classifier makes a prediction in the form of a statistical likelihood of a sample belonging to a particular class \cite{Goodfellow,James}. A $k$- categories classifier contains $k$ output neurons, each one representing a different class. When a set of features is presented to and transformed by the neural network, the value contained within each output neuron is the "confidence" on the interval $[0,1]$ that a sample belongs to that class. For example, if neuron $q$ has an activation value of $0.25$, we say that there is a $25\%$ chance that the sample belongs to class $q$. Most commonly, we take the neuron with the highest activation value to be the chosen class, even if the next highest activation is a close second \cite{Loy}. Thus the formal prediction function is the index of the maximum value in the output vector.

\paragraph*{}This concept stems from the idea that when a trained neural network classifier makes a prediction, it is basing class probabilities on a set of generated \textit{decision boundaries} \cite{Geron,James}. To explore decision boundaries, consider a simplified toy data set with a 2D feature space for a classifier that must only discern between \textit{Class A} and \textit{Class B}. We can visualize each sample in this 2D space, and color-code each sample according to its class label as in Fig. (\ref{fig-LinSep1}).

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.35]{../MakeFigures/LinearNonSeparable}
\end{center}
\caption{Feature Space with linear non-separability between two classes}
\label{fig-LinNonSep}
\end{figure}

@@

\item[•]University of New Hampshire College of Engineering and Physical Sciences, Department of Physics and Astronomy, and Department of Mathematics and Statistics.
